---
title: "Shiny EndomineR"
subtitle: "Refactoring as a way of life"
author: "Chris Beeley"
institute: "Nottinghamshire Healthcare NHS Trust"
date: "17th November 2022"
output:
  xaringan::moon_reader:
    css: 
      - default
      - css/nhsr.css
      - css/nhsr-fonts.css
    lib_dir: libs
    seal: false
    nature:
      highlightStyle: googlecode
      highlightLines: true
      highlightLanguage: ["r"]
      countIncrementalSlides: false
      ratio: "16:9"
    includes:
      after_body: [css/insert-logo.html]
---

```{r setup, include = FALSE}
library(knitr)
library(tidyverse)
library(NHSRtheme)
# set default options
opts_chunk$set(echo = FALSE)#,
               # fig.width = 7.252,
               # fig.height = 4,
               # dpi = 300)

# uncomment the following lines if you want to use the NHS-R theme colours by default
# scale_fill_continuous <- partial(scale_fill_nhs, discrete = FALSE)
# scale_fill_discrete <- partial(scale_fill_nhs, discrete = TRUE)
# scale_colour_continuous <- partial(scale_colour_nhs, discrete = FALSE)
# scale_colour_discrete <- partial(scale_colour_nhs, discrete = TRUE)
```

class: title-slide, left, bottom

# `r rmarkdown::metadata$title`
----
## **`r rmarkdown::metadata$subtitle`**
### `r rmarkdown::metadata$author`
### `r rmarkdown::metadata$date`

---

# Summary

* EndomineR
* The original Shiny EndomineR
* Refactoring to {golem}
* Refactoring again (?)

---

# EndomineR

* [Text mining for clinical text](https://github.com/ropensci/EndoMineR)
* Solves three main problems
  * General clean up, processing, spell check etc.
  * Data linkage
  * Analysis of linked data (e.g. looking at dosing of drugs)
  
---
  
# Shiny EndomineR

* This is a pretty typical Shiny application
* No modularisation
* Business logic in server.R code
* Data model is actually a little complicated

---

# Diagram

```{r}

DiagrammeR::mermaid("
graph TD
  A[Import endoscopy data]-->B[Import pathology data]
  B-->C[Merge data]
  C-->D[Map terms]
  D-->E[Save mapping]
  D-->F[Barrett's]
  D-->G[Per endoscopist]
  D-->H[Etc...]
")

```

---

# No brainer for golem!

* Package dependencies!
  * Use `::` (insert horror story here)
* Modularise the data reads to control the flow through the application
* Modularise the analytic code to make it reliable and flexible

---

# [“Stratégie du petit r”](https://engineering-shiny.org/structuring-project.html)

```{r echo = TRUE, eval = FALSE}

mod_clean_and_merge_server("clean_and_merge_ui_1",
                           header_filename = "endo.rda",
                           r = r)

mod_clean_and_merge_server("clean_and_merge_ui_2",
                           header_filename = "path.rda",
                           r = r)

mod_merge_data_server("merge_data_ui_1",
                      load_prev = load_prev,
                      r = r)
```

---

# More server.R

```{r echo = TRUE, eval = FALSE}

mod_map_terms_server("map_terms_ui_1", r = r)

barretts_data <- mod_barretts_server("barretts_ui_1", r = r)

polyp_data <- mod_polyps_server("polyps_ui_1", r = r)

mod_per_endoscopist_server("per_endoscopist_ui_1", 
                           barretts_data = barretts_data,
                           polyp_data = polyp_data, 
                           r = r)

mod_custom_server("custom_ui_1", r = r)

```

---

# Refactoring the business logic

* [Hope this works again!](https://github.com/nhs-r-community/shinyEndomineR/pull/49/files)

---

# Tips and tricks

* Everything should be as small as possible- but no smaller
* Write tests that test the business logic and the Shiny
* Use GitHub!
* Build outside of a reactive database

---

# Next level Shiny clinical applications

* Build end to end as soon as possible
* Don't be afraid to push your R/functions back into the business package they call
* Clinical datasets are really big and complicated and annoying
  * It may help the Shiny make sense to write more code to manipulate the raw data
